---
title: "Automatic or Manual - Which is Better for MPG?"
author: "Carlos Schuler"
date: "8/17/2020"
output: 
  ioslides_presentation:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Automatic or manual better for MPG??

### A quick look at the *mtcars* dataset:

```{r intialize, include=FALSE}
library(tidyverse)
library(plotly)
data("mtcars")
mtcars <- as_tibble(mtcars)
mtcarsF<-mtcars
mtcarsF$cyl <-  factor(mtcarsF$cyl)
mtcarsF$vs <-   factor(mtcarsF$vs,labels = c("V block","S block"))
mtcarsF$am <-   factor(mtcarsF$am,labels = c("Automatic","Manual"))
mtcarsF$gear <- factor(mtcarsF$gear)
mtcarsF$carb <- factor(mtcarsF$carb)
wtMedian <- median(mtcarsF$wt)
lightWeight=paste("<",wtMedian)
heavyWeight=paste(">",wtMedian)
mtcarsF <- mtcarsF %>% mutate(wtQ=factor(ntile(wt,2),labels=c(lightWeight,heavyWeight)))
rownames(mtcarsF)<-rownames(mtcars)
```
```{r mpgPlot, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", fig.height=3}
ax <- list(zeroline = TRUE, showline = TRUE, mirror = "ticks", gridcolor = "#F2F2F2",
  gridwidth = 0, zerolinecolor = toRGB("red"), zerolinewidth = 1, linecolor = toRGB("black"),
  linewidth = 2
)
ay <- list(zeroline = TRUE, showline = TRUE, mirror = "ticks", gridcolor = toRGB("lightblue"),
  gridwidth = 1, zerolinecolor = toRGB("red"), zerolinewidth = 1, linecolor = toRGB("black"),
  linewidth = 2
)
y1<-mtcarsF[mtcarsF$am=="Automatic",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic") %>% 
        add_trace(y= y2, name="Manual") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig
```

### suggests that **Manual transmissions** yield a better MPG

### BUT ...

## Grouping by Number of Cylinders ...

```{r mpgPlot2-1, echo=FALSE,warning=FALSE, message=FALSE,fig.align="center", fig.height=1.9}
y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="4",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="4",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 4 cyl") %>% 
        add_trace(y= y2, name="Manual, 4 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig

y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="6",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="6",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 6 cyl") %>% 
        add_trace(y= y2, name="Manual, 6 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig

y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="8",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="8",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 8 cyl") %>% 
        add_trace(y= y2, name="Manual, 8 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig
```

## Grouping by Weight ...

```{r mpgPlot2-2, echo=FALSE,warning=FALSE, message=FALSE,fig.align="center", fig.height=1.9}
y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$wtQ==lightWeight,]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$wtQ==lightWeight,]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, Light") %>% 
        add_trace(y= y2, name="Manual, Light") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig

y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$wtQ==heavyWeight,]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$wtQ==heavyWeight,]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, Heavy") %>% 
        add_trace(y= y2, name="Manual, Heavy") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',xaxis=ax, yaxis=ay)
fig
```

## We notice it is not that clear!!

- For 4-cyl vehicles, a manual transmission appears to yield higher MPG
- However, that appears **not to be the case** for 8-cyl vehicles 
- Also, for lighter vehicles (weighing less than the median for the sample), the relationship between MPG and transmission type appears not to be significant 
    + (*note*: there is only one vehicle weighing higher than the median weight with an manual transmission in the sample, the **Maserati Bora**. Its MPG does not appear to be better than other vehicles in its weight class.)

## CODE
```{r showcode, echo=TRUE, eval=FALSE}
library(tidyverse)
library(plotly)
data("mtcars")
mtcars <- as_tibble(mtcars)
mtcarsF<-mtcars
mtcarsF$cyl <-  factor(mtcarsF$cyl)
mtcarsF$vs <-   factor(mtcarsF$vs,labels = c("V block","S block"))
mtcarsF$am <-   factor(mtcarsF$am,labels = c("Automatic","Manual"))
mtcarsF$gear <- factor(mtcarsF$gear)
mtcarsF$carb <- factor(mtcarsF$carb)
wtMedian <- median(mtcarsF$wt)
lightWeight=paste("<",wtMedian)
heavyWeight=paste(">",wtMedian)
mtcarsF <- mtcarsF %>% mutate(wtQ=factor(ntile(wt,2),
                        labels=c(lightWeight,heavyWeight)))
rownames(mtcarsF)<-rownames(mtcars)
```

***
```{r showcode2, echo=TRUE, eval=FALSE}
ax <- list(zeroline = TRUE, showline = TRUE, mirror = "ticks", 
  gridcolor = "#F2F2F2", gridwidth = 0, 
  zerolinecolor = toRGB("red"), zerolinewidth = 1, 
  linecolor = toRGB("black"), linewidth = 2)
ay <- list(zeroline = TRUE, showline = TRUE, mirror = "ticks", 
  gridcolor = toRGB("lightblue"), gridwidth = 1, 
  zerolinecolor = toRGB("red"), zerolinewidth = 1, 
  linecolor = toRGB("black"), linewidth = 2)

y1<-mtcarsF[mtcarsF$am=="Automatic",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic") %>% 
        add_trace(y= y2, name="Manual") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2'
               ,xaxis=ax, yaxis=ay)
fig
```

***
```{r showcode3, echo=TRUE, eval=FALSE}
y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="4",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="4",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 4 cyl") %>% 
        add_trace(y= y2, name="Manual, 4 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',
               xaxis=ax, yaxis=ay)
fig

y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="6",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="6",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 6 cyl") %>% 
        add_trace(y= y2, name="Manual, 6 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',
               xaxis=ax, yaxis=ay)
fig
```

***
```{r showcode4, echo=TRUE, eval=FALSE}
y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$cyl=="8",]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$cyl=="8",]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, 8 cyl") %>% 
        add_trace(y= y2, name="Manual, 8 cyl") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',
               xaxis=ax, yaxis=ay)
fig

y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$wtQ==lightWeight,]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$wtQ==lightWeight,]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, Light") %>% 
        add_trace(y= y2, name="Manual, Light") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',
               xaxis=ax, yaxis=ay)
fig
```

***
```{r showcode5, echo=TRUE, eval=FALSE}
y1<-mtcarsF[mtcarsF$am=="Automatic" & mtcarsF$wtQ==heavyWeight,]$mpg
y2<-mtcarsF[mtcarsF$am=="Manual"    & mtcarsF$wtQ==heavyWeight,]$mpg
fig <- plot_ly(y= y1, type="box", name="Automatic, Heavy") %>% 
        add_trace(y= y2, name="Manual, Heavy") %>% 
        layout(paper_bgcolor='#F2F2F2', plot_bgcolor='#F2F2F2',
               xaxis=ax, yaxis=ay)
fig
```